---
title: "p8105_hw2_sz3214"
author: "Shizhe Zhang"
date: "2024-09-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

## Import and Clean data
```{r}
mta = 
  read_csv('data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv', na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(line:entry, vending, ada) |> 
  mutate(entry=ifelse(entry == "YES", TRUE, FALSE))
head(mta, 10)
```

The dataset contains variables: `r names(mta)`. 
First, read the cvs file and clean its column names.
Second, select all the variables you want.
Third, change the `entry` variable from character to logical.
The dataset has `r nrow(mta)` rows and `r ncol(mta)` columns.
The dataset is not tidy enough.

## Answer questions

```{r}
distinct(mta, line, station_name) |> 
  nrow()
```
There are 465 distinct stations.

```{r}
filter(mta, ada == TRUE) |> 
  distinct(line, station_name) |> 
  nrow()
```
There are 84 ADA compliant.

```{r}
filter(mta, entry = TRUE, vending == "NO") |> 
  summarise(proportion = n() / nrow(mta)) 
```
The proportion of station entrances / exits without vending allow entrance is 9.8%.


```{r}
mta_edt = mta |> 
  mutate(across(route1:route11, as.character)) |> 
  pivot_longer(
    cols = route1:route11, names_to = "route_number", values_to = "route") 

mta_edt |> 
  filter(route == "A") |> 
  distinct(line, station_name) |> 
  count()

mta_edt |> 
  filter(route == "A", ada == TRUE) |> 
  distinct(line, station_name) |> 
  count()
```
80 distinct stations serve the A train. 
Of the stations that serve the A train, 17 are ADA compliant.


## Problem 2

```{r}
mr_trash =
  read_excel('data/202409 Trash Wheel Collection Data.xlsx', sheet = 1, na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  mutate(sports_balls = as.integer(round(sports_balls)),
         trash_type = "Mr. Trash Wheel", 
         year = as.integer(year))

pro_trash =
  read_excel('data/202409 Trash Wheel Collection Data.xlsx', sheet = 2, na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  mutate(trash_type = "Professor Trash Wheel", 
         year = as.integer(year))

gwynnda=
  read_excel('data/202409 Trash Wheel Collection Data.xlsx', sheet = 4, na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  select(dumpster:homes_powered) |> 
  mutate(trash_type = "Gwynnda Trash Wheel", 
         year = as.integer(year))

trash = 
  bind_rows(mr_trash, pro_trash, gwynnda)  
```

There are `r nrow(trash)` of observations in the trash dataset and variables including `r names(trash)`.

```{r}
pro_trash |> 
  summarise(sum(weight_tons, na.rm = TRUE))
```

Total weifht of trash collected by Professor Trash Wheel is 488 tons.

```{r}
gwynnda |> 
  filter(year==2022, month=="June") |> 
  summarise(sum(cigarette_butts, na.rm = TRUE))
```

The total number of cigarette butts collected by Gwynnda in June of 2022 is 18120.

# Problem 3

```{r}
bakers = 
  read_csv('data/gbb_datasets/bakers.csv', na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  separate(baker_name, into = c("first_name", "last_name"), sep = " ") 

bakes = 
  read_csv('data/gbb_datasets/bakes.csv', na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  rename(first_name=baker)

results = 
  read_csv('data/gbb_datasets/results.csv', skip=2, na = c("NA", ".", "")) |> 
  janitor::clean_names() |> 
  rename(first_name=baker)
  
viewers = 
  read_csv('data/gbb_datasets/viewers.csv', na = c("NA", ".", "")) |>
  janitor::clean_names() |> 
  pivot_longer(
    cols = series_1:series_10,
    names_to = "series",
    values_to = "viewer"
  ) |> 
  mutate(series = str_replace(series, "series_", "")) |> 
  mutate(series = as.integer(series))
```

check and correct tables

```{r}
anti_join(bakers, bakes, by = "first_name")
anti_join(bakers, results, by = "first_name")
anti_join(bakes, results, by = c("series", "episode"))

bakers|>
  mutate(first_name = ifelse(first_name == "Jo","Joanne", first_name))
```

merge data

```{r}
gbb = 
  left_join(bakes, bakers, by = c("first_name","series")) |> 
  left_join(results, by = c("first_name","series", "episode")) |> 
  left_join(viewers, by = c("series", "episode")) |> 
  relocate(series,episode,first_name,last_name, signature_bake,show_stopper,technical,result,baker_age, baker_occupation, hometown)
```

export

```{r}
write_csv(gbb, 'data/gbb_datasets/gbb.csv')
```
